Spin canonical divisors of log stacky curves

### 2015-06-24 stacky_write_up/spin_cring.tex
fixed all bad references

\begin{lem}
\label{lem:deg1_sat_ind}
Let $X$ be a genus $g$ curve with a $\mathbf{Q}$-divisor $D$ (as
defined by O'Dorney \todo{reference this-Section 1 of Evan's paper, or preferably define it in the background section})
satisfying $h^0(X, \lfloor{D}\rfloor)\ge 1$.  Suppose $P$ is not a base-point of $dD$ for all $d\in \mathbf{N}^+$ meaning that $R_D$ has generically chosen generators (in the appropriate respective degrees)
$u, x_1, \ldots, _m$ all with $\ord_P=0$ and such that $\deg(u)=1$.  Let $D$ be generated by $u, x_1, \ldots, x_m$ in degree at most $\bar{d}$ with relations generated 
in degree at most $\bar{\tau}$, where all $x_i$s also satisfy $\ord_P(u)(x_i)$.  Suppose $D' = D + \frac{\alpha}{\beta} P$
for some $\alpha,\beta \in \mathbf{N}$ such that
\begin{equation}
\label{eqn:deg1_sat_ind_dim}
	h^0(X, \lfloor{dD + d\frac{\alpha}{\beta} P \rfloor}) = h^0(X,\lfloor dD
	\rfloor) + \lfloor d\frac{\alpha} {\beta} \rfloor \; \forall d \in \mathbf{
	N} : d \ge \frac{\beta}{\alpha}.
\end{equation}
Let 
\[
	0<\frac{c_1}{d_1}<...<\frac{c_n}{d_n}=\frac{\alpha}{\beta}
\]
be the non-negative best lower approximation.
Then 
\begin{itemize}
\item $R_{D'}$ is generated over $R_D$ by elements $y_1,...,y_n$ such that $y_i$ lies in  
degree $d_i$ and has a pole at $P$ of degree $c_i$

@@

In a fashion analogous to the proof of Lemma 
~\ref{lem:sat_three_induction}, we see 
the following is a $k$ basis for $R$ over $R'$:

@@

\iffalse{
\begin{lem}\label{lem:adding_points_saturation_2_to_1_with_log}
Let $(\sx, \Delta, \halfcan)$ and $(\sx, \Delta', L')$ be tame separably rooted log stacky spin 
curves with $0\ne \Delta=\Delta'+2P$ for some non-stacky point $P$ not in the support of $
\halfcan'$, meaning that $L=L'+P$.  Let $(g; e_1, \ldots, e_\tau; \delta+2)$ and $(g; e_1, 
\ldots, e_\tau, \delta)$ with $g\ge 2$ be their respective signatures.  Suppose 
$sat(Eff(\halfcan))=1$ and $sat(Eff(\halfcan'))=2$.  Further suppose $R_{\halfcan'} = k[x_2, 
x_3, \ldots, x_m]/I'$ with $\deg(x_2) = 2$ and $\deg(x_3) = 3$.  Then
\begin{enumerate}
\item[(a)] Then generic elements $y_1$ and $y_2$ in $(R_{\halfcan})_1$ and $
(R_{\halfcan})_2$ respectively, such that $-\ord_P(y_2)>0$ and $\{y_1^2, y_2\}$ is linearly 
independent, generate $R_\halfcan$ over $R_{\halfcan'}$.
\item[(b)] Equip $k[y_1,y_2]$ with $grlex$ so that $y_2 \prec y_1$ and equip $k[y_1, y_2, 
x_2, \ldots, x_m]$ with the block order so that $R_\halfcan=k[y_1, y_2, x_2, \ldots, x_m]/I$.  
Then
\begin{align*}
			in_\prec(I) &= in_\prec(I')k[x_2, \ldots, x_m, y_1, y_2] \\
			&+\langle y_2 x_j \mid 2 \leq j \leq m \rangle \\
			&+\langle y_2^2\rangle.
		\end{align*}
\item[(c)] Any set of minimal generators for $I'$ together with any set of relations with 
leading terms as above minimally generate $I$.
\end{enumerate}
\end{lem}

@@

\section{High genus}
\label{sec:g_high}
We now consider the case when the genus is at least 2. In this case, we are able to bound the degrees and relations of the half canonical ring $R_L$. In this section, we do not obtain explicit presentations of $R_L.$ In contrast, in Section ~\ref{sec:g_1} and Subsection ~\ref{ssec:g_0_effective} we will not only obtain bounds on the generators and relations, but also obtain presentations inductively. The tradeoff is that in the cases of genus 0 and 1, we will have to explicitly deal with base cases, making the induction more involved. We begin with the case that $g \geq 2$.

@@

In the case of a smooth genus 1 curve, $X$, we know $K_X = 0$, and therefore the only possibilities for half canonical divisors are $\halfcan' = 0$ or $\halfcan' = P-Q$ where $P,Q$ are distinct points, fixed under the hyperelliptic involution. We aim to inductively construct presentations by increasing the values of the $e_i$ using Theorem ~\ref{thm:ram_order_ind} and add additional points using Lemma ~\ref{lem:deg1_sat_ind} and Proposition ~\ref{lem:sat_two_induction}.

@@

There are two main cases: when $\delta \geq 2$ and when $\delta = 0$
 (note that $\delta$ must be odd or zero because $\halfcan \in \text
{ Div } \sx$). In the first case, $\deg \lfloor \halfcan \rfloor
\geq 0$ so the result directly follows from Lemma
~\ref{rem:deg1_sat_ind}. When $\delta = 0$, this is not necessarily
true so the proof is a rather involved induction using Lemma
~\ref{lem:sat_three_ induction} and Theorem ~\ref{thm:ram_order_ind}.
The remainder of this section is dedicated to characterizing the
saturations ( Subsection ~\ref{ssec:g_0_saturation}), inductive
base cases ( Subsection ~\ref{ssec:g_0_base}), and exceptional
cases (Subsection ~\ref{ssec:g_0_exceptional}) when $\delta = 0$
before proving the main theorem in full for all $\delta$ in
Subsection ~\ref{ssec:g_0_ main}.

@@

Through the application of two processes of induction, by adding 
points via Lemma ~\ref{sat-three-induction-g-0} and by raising
stabilizer orders via Theorem ~\ref{thm:ram_order_ind},
we obtain Theorem ~\ref{thm:g_0_generators_relations} for any tame
log spin curve $(\sx, \Delta, L)$ over a perfect field $k$ with
signature $\sigma$ covered by this induction process.
In particular, this induction process covers all signatures $\sigma
= (0; e_1, \ldots, e_r; 0)$ such that either $r \geq 5$ or
for some listed base case $(\sigma', J)$ with $\sigma' = (0; e_1',
\ldots, e_r'; 0)$, $e_i \geq e_i'$ for all $i$.
Before we prove the main theorem, we describe the cases not be
covered by induction.

@@

Now let us consider the case when $\delta \geq 2$. In any such case,
$\lfloor \halfcan \rfloor$ is an effective divisor and the conditions
of Lemma ~\ref{rem:deg1_sat_ind} are satisfied. Thus, we can apply
the lemma inductively on the classical case with no stacky points
to get that $R(\sx, \Delta, \halfcan)$ is generated up to degree $e$
with relations generated up to degree $2e$.

@@

If $r > 5$, then we can use Lemma
~\ref{sat-three-induction-g-0} to add ramified points with
stabilizer order $3$ to case (k) of Lemma
~\ref{lem:g_0_admissible_cases}, which corresponds to $\sigma_{11}
= (0; 3, 3, 3, 3, 3; 0), J_{11} = \{1, 2, 3, 4, 5\}$. This case
clearly satisfies the conditions of Lemma
~\ref{sat-three-induction-g-0} (recall from Subsection ~\ref
{ssec:g_0_saturation} that saturation is 3 here), and the
immediate consequence of parts (a) and (c) of Lemma
~\ref{sat-three-induction-g-0} is that any $R(\sx', \Delta,
\halfcan')$ corresponding to signatures $\sigma'$ with ramification
orders all equal to $3$ for any $r > 5$ is generated up to
degree $e := \max(5, e_1', \ldots, e_r')$ with relations generated
up to degree $2e$. Furthermore, Lemma ~\ref{lem:n-3s-admissible}
implies that these cases satisfy all of the conditions of Theorem
~\ref{thm:ram_order_ind}.

@@

All such base cases satisfy all of the hypotheses of
Theorem ~\ref{thm:ram_order_ind} and their corresponding $R' :=
R(\sx', \Delta, \halfcan')$ are generated, as $k$-algebras, up
to degree $e' := \max(5, e_1', \ldots, e_r')$ with relations
generated up to degree $2e'$ by Lemma ~\ref{g_0_admissible_cases}.
It follows from Theorem ~\ref{thm:ram_order_ind}(b) that $R :=
R(\sx, \Delta, \halfcan)$ is generated up to degree $e := \max(5,
e_1, \ldots, e_r)$. It follows from Theorem ~\ref{thm:ram_order_ind}
(d) that the relations are generated up to degree $2e$.
\end{proof}

\section{Further Research}
\label{sec:further-questions}
\begin{itemize}
	\item More on fractional modular forms
	\item Make explicit generators and relations for base case of non-stacky higher genus half-canonical divisors
	\item In the case $g \geq 2$, can the bound from Proposition ~\ref{prop:relation_generation} be reduced from 11 to 10 (which would then be sharp by Remark ~\ref{rem:relations_generation_ten})? \todo{Possibly think more about this} 
	\item Generic Initial ideal (and other genericness properties?)
\end{itemize}

>>>>

\begin{lem}
\label{lem:deg1-sat-ind}
Let $X$ be a genus $g$ curve with a $\mathbf{Q}$-divisor $D$ (as
defined by O'Dorney \todo{reference this-Section 1 of Evan's paper, or preferably define it in the background section})
satisfying $h^0(X, \lfloor{D}\rfloor)\ge 1$.  Suppose $P$ is not a base-point of $dD$ for all $d\in \mathbf{N}^+$ meaning that $R_D$ has generically chosen generators (in the appropriate respective degrees)
$u, x_1, \ldots, _m$ all with $\ord_P=0$ and such that $\deg(u)=1$.  Let $D$ be generated by $u, x_1, \ldots, x_m$ in degree at most $\bar{d}$ with relations generated 
in degree at most $\bar{\tau}$, where all $x_i$s also satisfy $\ord_P(u)(x_i)$.  Suppose $D' = D + \frac{\alpha}{\beta} P$
for some $\alpha,\beta \in \mathbf{N}$ such that
\begin{equation}
\label{eqn:deg1_sat_ind_dim}
	h^0(X, \lfloor{dD + d\frac{\alpha}{\beta} P \rfloor}) = h^0(X,\lfloor dD
	\rfloor) + \lfloor d\frac{\alpha} {\beta} \rfloor \; \forall d \in \mathbf{
	N} : d \ge \frac{\beta}{\alpha}.
\end{equation}
Let 
\[
	0<\frac{c_1}{d_1}<...<\frac{c_n}{d_n}=\frac{\alpha}{\beta}
\]
be the non-negative best lower approximation.
Then 
\begin{itemize}
\item $R_{D'}$ is generated over $R_D$ by elements $y_1,...,y_n$ such that $y_i$ lies in  
degree $d_i$ and has a pole at $P$ of degree $c_i$

@@

In a fashion analogous to the proof of Lemma 
~\ref{lem:sat-three-induction-g-0}, we see 
the following is a $k$ basis for $R$ over $R'$:

@@

\iffalse{
\begin{lem}
\label{lem:adding_points_saturation_2_to_1_with_log}
Let $(\sx, \Delta, \halfcan)$ and $(\sx, \Delta', L')$ be tame separably rooted log stacky spin 
curves with $0\ne \Delta=\Delta'+2P$ for some non-stacky point $P$ not in the support of $
\halfcan'$, meaning that $L=L'+P$.  Let $(g; e_1, \ldots, e_\tau; \delta+2)$ and $(g; e_1, 
\ldots, e_\tau, \delta)$ with $g\ge 2$ be their respective signatures.  Suppose 
$sat(Eff(\halfcan))=1$ and $sat(Eff(\halfcan'))=2$.  Further suppose $R_{\halfcan'} = k[x_2, 
x_3, \ldots, x_m]/I'$ with $\deg(x_2) = 2$ and $\deg(x_3) = 3$.  Then
\begin{enumerate}
\item[(a)] Then generic elements $y_1$ and $y_2$ in $(R_{\halfcan})_1$ and $
(R_{\halfcan})_2$ respectively, such that $-\ord_P(y_2)>0$ and $\{y_1^2, y_2\}$ is linearly 
independent, generate $R_\halfcan$ over $R_{\halfcan'}$.
\item[(b)] Equip $k[y_1,y_2]$ with $grlex$ so that $y_2 \prec y_1$ and equip $k[y_1, y_2, 
x_2, \ldots, x_m]$ with the block order so that $R_\halfcan=k[y_1, y_2, x_2, \ldots, x_m]/I$.  
Then
\begin{align*}
			in_\prec(I) &= in_\prec(I')k[x_2, \ldots, x_m, y_1, y_2] \\
			&+\langle y_2 x_j \mid 2 \leq j \leq m \rangle \\
			&+\langle y_2^2\rangle.
		\end{align*}
\item[(c)] Any set of minimal generators for $I'$ together with any set of relations with 
leading terms as above minimally generate $I$.
\end{enumerate}
\end{lem}

@@

\section{High genus}
\label{sec:g_high}
We now consider the case when the genus is at least 2. In this case, we are able to bound the degrees and relations of the half canonical ring $R_L$. In this section, we do not obtain explicit presentations of $R_L.$ In contrast, in Section ~\ref{sec:g_1} and Section ~\ref{sec:g_0} we will not only obtain bounds on the generators and relations, but also obtain presentations inductively. The tradeoff is that in the cases of genus 0 and 1, we will have to explicitly deal with base cases, making the induction more involved. We begin with the case that $g \geq 2$.

@@

In the case of a smooth genus 1 curve, $X$, we know $K_X = 0$, and therefore the only possibilities for half canonical divisors are $\halfcan' = 0$ or $\halfcan' = P-Q$ where $P,Q$ are distinct points, fixed under the hyperelliptic involution. We aim to inductively construct presentations by increasing the values of the $e_i$ using Theorem ~\ref{thm:ram_order_ind} and add additional points using Lemma ~\ref{lem:deg1-sat-ind} and Lemma ~\ref{lem:sat-two-induction-high-genus}.

@@

There are two main cases: when $\delta \geq 2$ and when $\delta = 0$
 (note that $\delta$ must be odd or zero because $\halfcan \in \text
{ Div } \sx$). In the first case, $\deg \lfloor \halfcan \rfloor
\geq 0$ so the result directly follows from Lemma
~\ref{lem:deg1-sat-ind}. When $\delta = 0$, this is not necessarily
true so the proof is a rather involved induction using Lemma
~\ref{lem:sat-three-induction-g-0} and Theorem ~\ref{thm:ram_order_ind}.
The remainder of this section is dedicated to characterizing the
saturations (Subsection ~\ref{ssec:g_0_saturation}), inductive
base cases (Subsection ~\ref{ssec:g_0_base}), and exceptional
cases (Subsection ~\ref{ssec:g_0_exceptional}) when $\delta = 0$
before proving the main theorem in full for all $\delta$ in
Subsection ~\ref{ssec:g_0_main}.

@@

Through the application of two processes of induction, by adding 
points via Lemma ~\ref{lem:sat-three-induction-g-0} and by raising
stabilizer orders via Theorem ~\ref{thm:ram_order_ind},
we obtain Theorem ~\ref{thm:g_0_generators_relations} for any tame
log spin curve $(\sx, \Delta, L)$ over a perfect field $k$ with
signature $\sigma$ covered by this induction process.
In particular, this induction process covers all signatures $\sigma
= (0; e_1, \ldots, e_r; 0)$ such that either $r \geq 5$ or
for some listed base case $(\sigma', J)$ with $\sigma' = (0; e_1',
\ldots, e_r'; 0)$, $e_i \geq e_i'$ for all $i$.
Before we prove the main theorem, we describe the cases not be
covered by induction.

@@

Now let us consider the case when $\delta \geq 2$. In any such case,
$\lfloor \halfcan \rfloor$ is an effective divisor and the conditions
of Lemma ~\ref{lem:deg1-sat-ind} are satisfied. Thus, we can apply
the lemma inductively on the classical case with no stacky points
to get that $R(\sx, \Delta, \halfcan)$ is generated up to degree $e$
with relations generated up to degree $2e$.

@@

If $r > 5$, then we can use Lemma
~\ref{lem:sat-three-induction-g-0} to add ramified points with
stabilizer order $3$ to case (k) of Lemma
~\ref{lem:g_0_admissible_cases}, which corresponds to $\sigma_{11}
= (0; 3, 3, 3, 3, 3; 0), J_{11} = \{1, 2, 3, 4, 5\}$. This case
clearly satisfies the conditions of Lemma
~\ref{lem:sat-three-induction-g-0} (recall from Subsection ~\ref
{ssec:g_0_saturation} that saturation is 3 here), and the
immediate consequence of parts (a) and (c) of Lemma
~\ref{lem:sat-three-induction-g-0} is that any $R(\sx', \Delta,
\halfcan')$ corresponding to signatures $\sigma'$ with ramification
orders all equal to $3$ for any $r > 5$ is generated up to
degree $e := \max(5, e_1', \ldots, e_r')$ with relations generated
up to degree $2e$. Furthermore, Lemma ~\ref{lem:n-3s-admissible}
implies that these cases satisfy all of the conditions of Theorem
~\ref{thm:ram_order_ind}.

@@

All such base cases satisfy all of the hypotheses of
Theorem ~\ref{thm:ram_order_ind} and their corresponding $R' :=
R(\sx', \Delta, \halfcan')$ are generated, as $k$-algebras, up
to degree $e' := \max(5, e_1', \ldots, e_r')$ with relations
generated up to degree $2e'$ by Lemma ~\ref{lem:g_0_admissible_cases}.
It follows from Theorem ~\ref{thm:ram_order_ind}(b) that $R :=
R(\sx, \Delta, \halfcan)$ is generated up to degree $e := \max(5,
e_1, \ldots, e_r)$. It follows from Theorem ~\ref{thm:ram_order_ind}
(d) that the relations are generated up to degree $2e$.
\end{proof}

\section{Further Research}
\label{sec:further-questions}
\begin{itemize}
	\item More on fractional modular forms
	\item Make explicit generators and relations for base case of non-stacky higher genus half-canonical divisors
	\item In the case $g \geq 2$, can the bound from Proposition ~\ref{prop:relations_generation} be reduced from 11 to 10 (which would then be sharp by Remark ~\ref{rem:relations_generation_ten})? \todo{Possibly think more about this} 
	\item Generic Initial ideal (and other genericness properties?)
\end{itemize}
